Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model

### 2013-12-12 Paper/gammaRays.tex
Added affiliations and model overview graphics

\usepackage{commath}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{units}
\usepackage{hyperref}

@@

\author{Maxim S. Piskunov and Grigory I. Rubtsov}
\title{Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model}

@@

Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics. Recent observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} allowed for extensive study of these explosions \cite{Vianello:2013ela,Gehrels:2013xd}. These studies led to several interesting results. For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}. Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}. The very high energy radiation was discovered to be extended if compared to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}. Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we'll call them low and high energy bands). In this paper we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.

@@

>>>>

\usepackage{commath}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{subcaption}
\usepackage{units}
\usepackage{hyperref}

@@

\title{Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model}
\author{
	Maxim S. Piskunov$^{1}$\thanks{maxit@ms2.inr.ac.ru} and
	Grigory I. Rubtsov$^{1}$\thanks{grisha@ms2.inr.ac.ru} \\
	$^{1}$Institute for Nuclear Research RAS, 117312, Moscow, Russia\\
}

@@

Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics. Recent observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} allowed for extensive study of these explosions \cite{Vianello:2013ela,Gehrels:2013xd}. These and previous studies led to several interesting results. For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}. Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}. The very high energy radiation was discovered to be extended if compared to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}. Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we'll call them low and high energy bands). In this paper we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.

@@

We propose that this result can be explained by the curvature effects (that is the effects of the jet geometry). This effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}. However, these studies were only concerned with x-ray radiation, and, more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet. On the contrary, we propose a different idea: that the highest energy radiators are concentrated near the axis of the burst, that is the burst opening angle depends on energy.

\begin{figure}[modelOverview]
        \centering
        \includegraphics[width=\textwidth]{modelOverview}
        \caption{Model Overview}
        \label{fig:modelOverview}
\end{figure}
